---
title: "PermutationTests"
author: "Louie Alexander"
date: '2022-06-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#change access and/or change setwd for your computer
access = "Louie"

if(access =="Louie") {
  setwd("C:\\Users\\louie\\Desktop\\ResearchFiles\\RPlotting")
}

```

```{r}

if (!require(readxl)) {
  install.packages("readxl")
}
if (!require(tidyverse)) {
  install.packages("tidyverse")
}
if (!require(reshape2)) {
  install.packages("reshape2")
}

library(readxl)
library(tidyverse)
library(reshape2)

```
#Grab data and prepare dataframe
```{r}
duplications <- read.table("DuplicationsPerInterval_Example.csv",header=TRUE,sep = ',')

duplications <- duplications %>% select(-c(pi300119))

#Note that you need to shift counts for each strain to their own column, and have a column for strain name.
#Use gather() for this
tallData <- gather(duplications, strain, counts, rio:pi510757, factor_key = TRUE)
tallData

```

#Make a dataframe per chromosome for each of reference
```{r}
chr1 <- tallData %>% filter(CHROM == "Chr01") 
chr1 <- chr1[order(chr1$interval),]

chr2 <- tallData %>% filter(CHROM == "Chr02") 
chr2 <- chr2[order(chr2$interval),]

chr3 <- tallData %>% filter(CHROM == "Chr03") 
chr3 <- chr3[order(chr3$interval),]

chr4 <- tallData %>% filter(CHROM == "Chr04") 
chr4 <- chr4[order(chr4$interval),]

chr5 <- tallData %>% filter(CHROM == "Chr05") 
chr5 <- chr5[order(chr5$interval),]

chr6 <- tallData %>% filter(CHROM == "Chr06") 
chr6 <- chr6[order(chr6$interval),]

chr7 <- tallData %>% filter(CHROM == "Chr07") 
chr7 <- chr7[order(chr7$interval),]

chr8 <- tallData %>% filter(CHROM == "Chr08") 
chr8 <- chr8[order(chr8$interval),]

chr9 <- tallData %>% filter(CHROM == "Chr09") 
chr9 <- chr9[order(chr9$interval),]

chr10 <- tallData %>% filter(CHROM == "Chr10") 
chr10 <- chr10[order(chr10$interval),]
```

#make a function to get difference matix for each interval
```{r}
#writing function to return a difference matrix for a given dataframe above
getDifferences <- function(dataframe){

intervals <- length(dataframe$interval)/9
inter <- c(0:intervals)
diff <- c(rep(0,intervals+1))
diffMatrix <- data.frame(inter, diff)

for(i in 0:intervals){
  calc <- dataframe %>% filter(interval == i)
  diffMatrix[i+1,2] = abs(mean(calc$counts[calc$strain %in% c("rio", "leoti", "CA")]) - mean(calc$counts[calc$strain %in% c("pi297155", "pi229841", "pi329311", "pi506069", "pi655972", "pi510757")])) 
}
return(diffMatrix)
}
```

#test the function on chromosome 1 data
```{r}
chr1Diffs <- getDifferences(chr1)
```

#Make function for permutation tests that generates a p-value.
#Currently works for 1 interval of a chromosome
```{r}


#inter is the desired interval and tests is the desired number of tests
permutation.test <- function(dataframe, inters, tests){
  
differences <- getDifferences(dataframe)

dataframe <- dataframe %>% filter(interval == inters)

#set up a matrix with dimensions: number of strains x number of tests desired
permutationMatrix = matrix(0,nrow = 9, ncol = tests)

#initialize matrix for the test differences per permutation
testDiffs <- rep(0, tests)
  
  
for(i in 1:tests){
  permutationMatrix[, i] <- sample(dataframe$counts, size = 9, replace = FALSE)
}

for (i in 1:tests){
    testDiffs[i] <- abs(mean(permutationMatrix[dataframe$strain %in% c("rio", "leoti", "CA"), i]) - mean(permutationMatrix[dataframe$strain %in% c("pi297155", "pi229841", "pi329311", "pi506069", "pi655972", "pi510757"), i]))
}  

pvalue = sum(testDiffs >= differences$diff[differences$inter==inters])/tests

return(pvalue)
}
```

#Test that function gives p-value
```{r}
test <- permutation.test(chr1, 0, 1000000)
print(test)
```

#Write a function that goes through each interval for a chromosome
```{r}

#set seed means we have a predefined number generator for our samplings
set.seed(1979000000) 

AllIntervals <- function(dataframe, test){
  
len <- length(dataframe$interval)/9
Intervals <- c(0:len)
Chromosome <- rep(dataframe[1,"CHROM"],len+1)
P_Value <- rep(0,len+1) 

results = data.frame(Chromosome, Intervals, P_Value)

for (i in Intervals){
  if (i %in% dataframe$interval){
  results[i+1,"P_Value"] = permutation.test(dataframe, i, test)
  }
  else{
    results[i+1,"P_Value"] = 1.00
  }
}
return(results)
}

```

#test that function gives dataframe of results
```{r}
#max number of permutations is n!
#factorial(9) #=362880

#1000 iterations is fine
#resultsChr1 <- AllIntervals(chr1, 1000)
#resultsChr1_5k <- AllIntervals(chr1,5000)
#resultsChr1_10k <- AllIntervals(chr1,10000)

#this number of permutations surpasses the max number
#resultsChr1_100k <- AllIntervals(chr1,100000)
```

#Looking at the independence test of coin for reference
```{r}
library(coin)

testing <- chr1 %>% filter(interval==1) %>% mutate(category = ifelse(strain %in% c("rio", "leoti", "CA"), "sweet", "biomass"))

#gives similar result as 1k iterations
independence_test(counts~factor(category), data=testing)
```

#Filter to significant differences of Chr01
```{r}
#signifChr1 <- resultsChr1 %>% filter(P_Value < 0.05)
#signifChr1 <- left_join(signifChr1, chr1Diffs, by = c("Intervals" = "inter"))
#signifChr1 <- signifChr1[order(signifChr1$P_Value),]

#replicate this in a function:
getSignificance <- function(ChrData, tests){
differences <- getDifferences(ChrData)

results <- AllIntervals(ChrData, tests)

signif <- results %>% filter(P_Value < 0.05)
signif <- left_join(signif, differences, by = c("Intervals" = "inter"))
signif <- signif[order(signif$P_Value),]

return(signif)
}

```

#Significance of all chromosomes
```{r}
#set seed means we have a predefined number generator for our samplings
set.seed(1979000000) 

finalChr1 <- getSignificance(chr1, 1000)
finalChr2 <- getSignificance(chr2, 1000)
finalChr3 <- getSignificance(chr3, 1000)
finalChr4 <- getSignificance(chr4, 1000)
finalChr5 <- getSignificance(chr5, 1000)
finalChr6 <- getSignificance(chr6, 1000)
finalChr7 <- getSignificance(chr7, 1000)
finalChr8 <- getSignificance(chr8, 1000)
finalChr9 <- getSignificance(chr9, 1000)
finalChr10 <- getSignificance(chr10, 1000)

```

#merge final data
```{r}
merge <- full_join(finalChr1, finalChr2)
merge <- full_join(merge, finalChr3)
merge <- full_join(merge, finalChr4)
merge <- full_join(merge, finalChr5)
merge <- full_join(merge, finalChr6)
merge <- full_join(merge, finalChr7)
merge <- full_join(merge, finalChr8)
merge <- full_join(merge, finalChr9)
merge <- full_join(merge, finalChr10)
```

```{r}
write.csv(merge, "SignificantDuplications.csv")
```

